---
title: "data"
output: html_document
---

Import initial dataset
---------------------------------------
```{r}
set.seed(123)
library(ggplot2)
train = read.csv('../raw_data/train (1).csv')
dim(train)

test = read.csv('../raw_data/test (1).csv')
test$Survived = NA
dim(test)

full = rbind(train, test) 
dim(full)
#str(full) 
```



```{r}
print('Missing observation?')
colSums(is.na(full))
print('Empty data')
colSums(full=='')
print('number of unique values per column')
apply(full,2, function(x) length(unique(x)))
```

EMBARKED
------------------------------------------------------------------------

```{r}
table(full$Embarked)
# replace missing with mode
full$Embarked[full$Embarked==""]="S"
table(full$Embarked)
full$Embarked = factor(full$Embarked)
```

Other methods:
```{r}
full$Embarked <- replace(full$Embarked, which(is.na(full$Embarked)), 'S')
```

```{r}

ggplot(full[1:891,],aes(x = Embarked, fill=factor(Survived))) +
geom_bar() +
ggtitle("Embarked vs Survival Rate")+
xlab("Embarked") +
ylab("Total Count") +
labs(fill = "Survived") 
# I can definitely test for an interaction
```

```{r}
tapply(train$Survived,train$Embarked,mean) # used train 
```
Note that C is relatively higher

PCLASS
------------------------------------------------------------------------

```{r}
full$Pclass = factor(full$Pclass)
# consider normalizing as numeric
```


```{r}
###Visualize P class which is the best proxy for Rich and Poor  

ggplot(full[1:891,],aes(x = Pclass,fill=factor(Survived))) +
geom_bar() +
ggtitle("Pclass v/s Survival Rate")+
xlab("Pclass") +
ylab("Total Count") +
labs(fill = "Survived")  
```

```{r}
# Visualize the 3-way relationship of sex, pclass, and survival
ggplot(full[1:891,], aes(x = Sex, fill = factor(Survived))) +
geom_bar() + #bar chart
facet_wrap(~Pclass) + 
ggtitle("3D view of sex, pclass, and survival") +
xlab("Sex") +
ylab("Total Count") +
labs(fill = "Survived")
```


```{r}
# Visualize the 3-way relationship of embarked, pclass, and survival
ggplot(full[1:891,], aes(x = Embarked, fill = factor(Survived))) +
geom_bar() + #bar chart
facet_wrap(~Pclass) + 
ggtitle("3D view of embarked, pclass, and survival") +
xlab("Embarked") +
ylab("Total Count") +
labs(fill = "Survived")
```



FARE
------------------------------------------------------------------------

```{r}

full$Fare = ifelse(is.na(full$Fare),
                     ave(full$Fare, FUN = function(x) mean(x, na.rm = TRUE)),
                     full$Fare)
# Just curious
plot(full$Fare, full$Age)
# there is less variance in age as the fare becomes higher
```





TITLE
------------------------------------------------------------------------
```{r}
#head(full$Name, n = 1)
names = full$Name
title =  gsub("^.*, (.*?)\\..*$", "\\1", names)
table(title)
```

If-else func for title assignment - preferred
```{r}
n = length(title)
for(i in 1:n){
  if(title[i] == "Mlle"){
    title[i] = "Miss"
  }else if(title[i] == "Ms"){
    title[i] = "Miss"
  }else if(title[i] == "Mme"){
    title[i] = "Mrs"
  }else if(title[i] == "Lady"){
    title[i] = "Miss"
  }else if(title[i] == "Dona"){
    title[i] = "Mrs"
  }else if(title[i] == "Don"){
    title[i] = "Master"
  }else if(title[i] == "Capt"){
    title[i] = "crew"
  }else if(title[i] == "Col"){
    title[i] = "crew"
  }else if(title[i] == "Major"){
    title[i] = "crew"
  }else if(title[i] == "Rev"){
    title[i] = "crew"
  }else if(title[i] == "Dr"){
    title[i] = "crew"
  }else if(title[i] == "Jonkheer"){
    title[i] = "Master"
  }else if(title[i] == "Sir"){
    title[i] = "Master"
  }else if(title[i] == "the Countess"){
    title[i] = "Mrs"
  }
}

full$title = factor(title)
table(full$title)
#str(full)
contrasts(full$title)
```



```{r}
ggplot(full[1:891,],aes(x = title, fill=factor(Survived))) +
geom_bar() +
ggtitle("Title vs Survival Rate")+
xlab("Title") +
ylab("Total Count") +
labs(fill = "Survived") 
```


```{r}
tapply(full$Survived[1:891], full$title[1:891], mean) 
```

```{r}
# Visualize the 3-way relationship of title, pclass, and survival
ggplot(full[1:891,], aes(x = title, fill = factor(Survived))) +
geom_bar() + #bar chart
facet_wrap(~Pclass) + 
ggtitle("3D view of title, pclass, and survival") +
xlab("title") +
ylab("Total Count") +
labs(fill = "Survived")
```


Alternate method - title
```{r}
full$title[full$title == 'Mlle'] = 'Miss' 
full$title[full$title == 'Ms']  = 'Miss'
full$title[full$title == 'Mme'] = 'Mrs' 
full$title[full$title == 'Lady'] = 'Miss'
full$title[full$title == 'Dona'] = 'Miss'
full$title[full$title == 'Capt'] = 'Officer' 
full$title[full$title == 'Col'] = 'Officer' 
full$title[full$title == 'Major'] = 'Officer'
full$title[full$title == 'Dr']  = 'Officer'
full$title[full$title == 'Rev'] = 'Officer'
full$title[full$title == 'Don'] = 'Officer'
full$title[full$title == 'Sir'] = 'Officer'
full$title[full$title == 'the Countess']  = 'Officer'
full$title[full$title == 'Jonkheer']  = 'Officer'

table(full$title)
```

FAMILY
------------------------------------------------------------------------
Primary
```{r}
sib = full$SibSp
pat = full$Parch
family = sib + pat + 1
n = length(family)
for(i in 1:n){
  if(family[i] == 1){
    family[i] = "solo"
  }else if(family[i] >= 2 & family[i] < 5){
    family[i] = "small"
  }else{
    family[i] = "large"
  }
}

table(family)
full$family = factor(family) 

```

Alternate method - Family
```{r}
# Lets create a Family size 

full_titanic$FamilySize = full_titanic$SibSp + full_titanic$Parch + 1

full_titanic$FamilySized[full_titanic$FamilySize == 1]   = 'Single'
full_titanic$FamilySized[full_titanic$FamilySize < 5 & full_titanic$FamilySize >= 2]   = 'Small'
full_titanic$FamilySized[full_titanic$FamilySize >= 5]   = 'Large'

full_titanic$FamilySized=as.factor(full_titanic$FamilySized)
```

Consider assigning converting to normalized numeric



AGE
------------------------------------------------------------------------
You may have to wait on this - and study up on decision trees
```{r}
library(xgboost)
library(rpart)
library(Ckmeans.1d.dp)
```

```{r}
dim(full[!is.na(full$Age),])
full[!is.na(full$Age), ]
names(full)
```

Recursive Partitioning and Regression Trees

```{r}
Agefit = rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + title + family, data=full[!is.na(full$Age),], method="anova")
        
full$Age[is.na(full$Age)] = predict(Agefit, full[is.na(full$Age),])
```


```{r}
# number of survivals and nonsurvivals across different age

ggplot(full[1:891,], aes(Age,fill = factor(Survived))) +
    geom_histogram()
```


Alternative method - average age
```{r}
full$Age = ifelse(is.na(full$Age),
                          ave(full$Age, FUN = function(x) mean(x, na.rm = TRUE)),
                          full$Age)
#na.rm: a logical value indicating whether NA values should be stripped before the computation proceeds.
```


SEX
------------------------------------------------------------------------

```{r}
ggplot(full[1:891,],aes(x = Sex, fill=factor(Survived))) +
geom_histogram(stat = "count")
#ggtitle("Sex vs Survival Rate")+
#xlab("Sex") +
#ylab("Total Count") +
#labs(fill = "Survived") 
```

CABIN
------------------------------------------------------------------------

```{r}

# Process Cabin Column to show number of cabins passenger has
cabin = as.character(full$Cabin)
n = length(cabin)
for(i in 1:n){
  if(nchar(cabin[i]) == 0){
    cabin[i] = 0
  } else{
    cabin[i] = length(strsplit(cabin[i]," ")[[1]])
  }
} 
table(cabin)
#looks good, use it for now
full$Cabin = factor(cabin)
str(full)
#consider numeric
```


```{r}

ggplot(full[1:891,],aes(x = Cabin,fill=factor(Survived))) +
geom_bar() +
ggtitle("Cabin v/s Survival Rate")+
xlab("Cabin") +
ylab("Total Count") +
labs(fill = "Survived")  
```



TICKET
------------------------------------------------------------------------

```{r}
t_unique = rep(0, nrow(full)) #zero array used to construct new column
tickets = unique(full$Ticket)

str(tickets) #929 unique ticket values
str(full$Ticket) #1309
str(t_unique) #1309 zero vector
n = length(tickets)
for(i in 1:n){
  group = which(full$Ticket == tickets[i]) #integer row vector consisting of index numbers where this is TRUE
  m = length(group)
  
  for(k in 1:m){
    t_unique[group[k]] = m
  }
}


```



```{r}
t_size = c()
n = length(t_unique)
for(i in 1:n){
  if(t_unique[i] == 1){
    t_size[i] = "solo"
  }else if(t_unique[i] >= 2 & t_unique[i] < 5){
    t_size[i] = "small"
  }else{
    t_size[i] = "large" 
  }
}

str(t_size)  
table(t_size)
full$ticket_group = factor(t_size)
# consider numeric
```


```{r}
ggplot(full[1:891,],aes(x = ticket_group,fill=factor(Survived))) +
geom_bar() +
ggtitle("Ticket v/s Survival Rate")+
xlab("Ticket") +
ylab("Total Count") +
labs(fill = "Survived")  
```

This could be an interaction between having muliple tickets and being a male. Assume most of the poor families lost the father/husband

Export to csv
By default, R uses traditional dummy coding (called “treatment contrasts” in R) for any non-ordered factors,
```{r}
ex = full[c("Pclass","Sex","Embarked", "family", "title", "Age", "Cabin", "ticket_group", "Fare", "Survived")]

# numerics are not normalized btw

library(caret)
dmy = dummyVars(" ~ .", data = ex, fullRank=T)
ex.new = data.frame(predict(dmy, newdata = ex))
#dummy trap solved

ex.new$Age = scale(as.numeric(ex.new$Age))
ex.new$Fare = scale(as.numeric(ex.new$Fare))

# Write CSV in R
write.csv(ex.new, file = "../processed_data/5_15.csv") #may have to manually drop dummy var
```




MATRIX
------------------------------------------------------------------------

```{r}

#ready to go for test
y = full[891:1309, c("Pclass","Sex","Embarked", "family", "title", "Age", "Cabin", "ticket_group", "Fare")]

X = full[1:891, c("Pclass","Sex","Embarked", "family", "title", "Age", "Cabin", "ticket_group", "Fare", "Survived")]

X$Age = scale(as.numeric(X$Age))
X$Fare = scale(as.numeric(X$Fare))

y$Age = scale(as.numeric(y$Age))
y$Fare = scale(as.numeric(y$Fare))
```




```{r}

library(caTools)

split = sample.split(X$Survived, SplitRatio = 0.75)
training_set = subset(X, split == TRUE)
test_set = subset(X, split == FALSE)

```





GLM
------------------------------------------------------------------------

```{r}
fit = glm(formula =  Survived ~ ., 
                 family = binomial,
                 data = training_set)

prob_pred = predict(fit, type = 'response', newdata = test_set)

head(prob_pred) # probability that dependent variable is equal to one
y_pred = ifelse(prob_pred > 0.5, 1, 0)
#y_pred[1:10]
#prob_pred[1:10]

# test_set[ , 10]
# Making the Confusion Matrix/accuracy
cm = as.matrix(table(Actual = test_set[ , 10], Predicted = y_pred))
paste("Accuracy = ", sum(diag(cm))/sum(cm))
```



KERNEL SVM
------------------------------------------------------------------------

```{r}
library(e1071)

fit = svm(formula =  Survived ~ ., data = training_set, 
          type = 'C-classification', 
          kernel = 'radial')
          
                 

prob_pred = predict(fit, type = 'response', newdata = test_set)

head(prob_pred) # don't need to write script for SVM 


# test_set[ , 10]
# Making the Confusion Matrix/accuracy
cm = as.matrix(table(Actual = test_set[ , 10], Predicted = prob_pred))
paste("Accuracy = ", sum(diag(cm))/sum(cm))
```



NAIVE BAYES
------------------------------------------------------------------------

```{r}

fit = naiveBayes(x = training_set[-10],
                        y = training_set$Survived)

prob_pred = predict(fit, newdata = test_set) #binary output

cm = as.matrix(table(Actual = test_set[ , 10], Predicted = prob_pred))
paste("Accuracy = ", sum(diag(cm))/sum(cm))
```



