---
title: "R preprocessing "
output: html_notebook
---


```{r}
library('ggplot2') 
library('caret') 
library('dplyr') 
library('car')

```

```{r}
train = read.csv('train (1).csv', stringsAsFactors=F)
test = read.csv('test (1).csv', stringsAsFactors = F)
test$Survived = NA
full = rbind(train, test) 
```


experiment rexp
Exploratory 

head: Returns the first or last parts of a vector, matrix, table, data frame or function. Since head() and tail() are generic functions, they may also have been extended to other classes.

gsub: sub and gsub perform replacement of the first and all matches respectively.
gsub(pattern, replacement, x, ignore.case = FALSE, perl = FALSE,
     fixed = FALSE, useBytes = FALSE)
     
regx: A ‘regular expression’ is a pattern that describes a set of strings. 

x <- "A. J. Burnett"
gsub("([A-Z])\\.\\s([A-Z])\\.", "\\1.\\2.", x)
[1] "A.J. Burnett"

this actually makes sense...

gsub('.*, |\\..*', '', combined_data$Name)
```{r}
test <- c("MAllen, Mr. William Henry", "Hewlett, Mrs. (Mary D Kingcome)", "Futrelle, Mrs. Jacques Heath (Lily May Peel)")

 gsub('(.*, )|(\\..*)', '', test)
```
Basically, replace (.*, ) everything before the comma and/or everything after the period (\\.) everything 

|    is OR operand
.*   Any character except line break
\\.  is a way to use period, otherwise . means "any character"

nchar: takes a character vector as an argument and returns a vector whose elements contain the sizes of the corresponding elements of x.

x <- c("asfef", "qwerty", "yuiop[", "b", "stuff.blah.yech")
nchar(x)
# 5  6  6  1 15

CABIN
```{r}
cabin = full$Cabin
dim(cabin)
summary(cabin)
str(cabin)
```

strsplit: Split the elements of a character vector x into substrings according to the matches to substring split within them.

```{r}
x = c(as = "asfef", qu = "qwerty", "yuiop[", "b", "stuff.blah.yech")
str(x)
# split x on the letter e
x1 = strsplit(x, "e") #this is now a list
typeof(x1)
str(x1) #still has the original number of elements; however, elements that have been split are now single element sublists

```


```{r}
cabin_t= cabin[1:50]
c = cabin_t[22] 
str(c)

s_t = strsplit(cabin_t[22], " ")
s_t #converts into a single element list, no actual split though

cabin_t[28]
s1 = strsplit(cabin_t[28], " ")[[1]] #tacking the double bracket on returns a clean 1 element list
#or a two step process
s2 = strsplit(cabin_t[28], " ")
s2[[1]]
length(s2[[1]])
```


```{r}

n = length(cabin_t) #50
for(i in 1:n){
  if(nchar(cabin_t[i]) == 0){
    cabin_t[i] = 0
  }else{
    cabin_t[i] = length(strsplit(cabin_t[i]," ")[[1]])
  }
} 
table(cabin_t)
```


```{r}
n = length(cabin_t)
for(i in 1:n){
  ifelse(nchar(cabin_t[i]) == 0, cabin_t[i] = 0, cabin_t[i] = length(strsplit(cabin_t[i]," ")[[1]]))}
   
table(cabin_t)
```
```{r}
cabin_t[28]
cabin_t[28] = length(strsplit(cabin_t[28]," ")[[1]])

ifelse(nchar(cabin_t[28]) == 0, cabin_t[28] = 0, cabin_t[28] = length(strsplit(cabin_t[28]," ")[[1]]))
#not sure why this func doesn't work

nchar(cabin_t[28]) == 0
cabin_t[28] = 0
cabin_t[28] = length(strsplit(cabin_t[28]," ")[[1]])

```

TITLES
```{r}
full$title[full$title == 'Mlle'] = 'Miss' 
full$title[full$title == 'Ms']  = 'Miss'
full$title[full$title == 'Mme'] = 'Mrs' 
full$title[full$title == 'Lady'] = 'Miss'
full$title[full$title == 'Dona'] = 'Miss'
full$title[full$title == 'Capt'] = 'Officer' 
full$title[full$title == 'Col'] = 'Officer' 
full$title[full$title == 'Major'] = 'Officer'
full$title[full$title == 'Dr']  = 'Officer'
full$title[full$title == 'Rev'] = 'Officer'
full$title[full$title == 'Don'] = 'Officer'
full$title[full$title == 'Sir'] = 'Officer'
full$title[full$title == 'the Countess']  = 'Officer'
full$title[full$title == 'Jonkheer']  = 'Officer'

table(full$title)
```


```{r}
names = full$Name
str(names)
title =  gsub("^.*, (.*?)\\..*$", "\\1", names)
str(title)
head(title, n = 10)
table(title)
```

```{r}
n = length(title)
for(i in 1:n){
  if(title[i] == "Mlle"){
    title[i] = "Miss"
  }else if(title[i] == "Ms"){
    title[i] = "Miss"
  }else if(title[i] == "Mme"){
    title[i] = "Mrs"
  }else if(title[i] == "Lady"){
    title[i] = "Miss"
  }else if(title[i] == "Dona"){
    title[i] = "Mrs"
  }else if(title[i] == "Don"){
    title[i] = "Master"
  }else if(title[i] == "Capt"){
    title[i] = "crew"
  }else if(title[i] == "Col"){
    title[i] = "crew"
  }else if(title[i] == "Major"){
    title[i] = "crew"
  }else if(title[i] == "Rev"){
    title[i] = "prof"
  }else if(title[i] == "Dr"){
    title[i] = "prof"
  }else if(title[i] == "Jonkheer"){
    title[i] = "Master"
  }else if(title[i] == "Sir"){
    title[i] = "Master"
  }else if(title[i] == "the Countess"){
    title[i] = "Mrs"
  }
}

table(title)



```

FAMILY
```{r}
# Lets create a Family size 

full_titanic$FamilySize = full_titanic$SibSp + full_titanic$Parch + 1

full_titanic$FamilySized[full_titanic$FamilySize == 1]   = 'Single'
full_titanic$FamilySized[full_titanic$FamilySize < 5 & full_titanic$FamilySize >= 2]   = 'Small'
full_titanic$FamilySized[full_titanic$FamilySize >= 5]   = 'Large'

full_titanic$FamilySized=as.factor(full_titanic$FamilySized)
```


```{r}
sib = full$SibSp
pat = full$Parch
family = sib + pat + 1
n = length(family)
for(i in 1:n){
  if(family[i] == 1){
    family[i] = "solo"
  }else if(family[i] >= 2 & family[i] < 5){
    family[i] = "small"
  }else{
    family[i] = "large"
  }
}

table(family)
```

TICKETS
```{r}
head(full$Ticket)
ticket = unique(full$Ticket)
length(ticket)

ticket_unique = rep(0, nrow(full))



```

when we're looking at tickets, we are looking for the same tickets. This will indicate families/groups?

```{r}
ticket_unique = rep(0, nrow(full)) #zero array used to construct new column
tickets = unique(full$Ticket)  #929 unique ticket values
head(tickets)
length(tickets)
for (i in 1:length(tickets)) {
  current_ticket = tickets[i]
  party_indexes = which(full$Ticket == current_ticket) #returns an integer row vector
  
  
  for (k in 1:length(party_indexes)) {
    ticket_unique[party_indexes[k]] = length(party_indexes)
  }
}
table(ticket_unique)


full$ticket_unique = ticket_unique #create a column


full$ticket_size[full$ticket_unique == 1]   <- 'Single'
full$ticket_size[full$ticket_unique < 5 & full$ticket_unique>= 2]   <- 'Small'
full$ticket_size[full$ticket_unique >= 5]   <- 'Big'



```

```{r}

test = full$Ticket[1:300]
test_unique = rep(0, 300) #creating a new column of zeros
#table(test)
#str(test)

#basically, the same idea as family as far as grouping
current_ticket = "382652"
index = which(test == current_ticket)
typeof(index)
str(index) #integer row vector
test[17]
index[1] #this equals 17
test_unique[index[1]] = length(index)

```


```{r}
ticket_unique = rep(0, nrow(full)) #zero array used to construct new column
tickets = unique(full$Ticket)

str(tickets) #929 unique ticket values
str(full$Ticket) #1309
str(ticket_unique) #1309 zero vector
n = length(tickets)
for(i in 1:n){
  group = which(full$Ticket == tickets[i]) #integer row vector consisting of index numbers where this is TRUE
  m = length(group)
  
  for(k in 1:m){
    ticket_unique[group[k]] = m
  }
}

str(ticket_unique)
table(ticket_unique)


```

```{r}
ticket_size = c()
n = length(ticket_unique)
for(i in 1:n){
  if(ticket_unique[i] == 1){
    ticket_size[i] = "solo"
  }else if(ticket_unique[i] >= 2 & ticket_unique[i] < 5){
    ticket_size[i] = "small"
  }else{
    ticket_size[i] = "large" 
  }
}

str(ticket_size)  
table(ticket_size)
full$ticket_group = ticket_size

```

```{r}
full$Age = ifelse(is.na(full$Age),
                          ave(full$Age, FUN = function(x) mean(x, na.rm = TRUE)),
                          full$Age)
```

My code




