---
title: "titanic"
output: html_document
---

Using Kernal 'Beginners Titanic'

Error message: The message means that both the packages have functions with the same names. In this particular case, the testthat and assertive packages contain five functions with the same name.
```{r}
library('ggplot2') 
library('dplyr') 
library('car')
```

Factor variables are categorical variables that can be either numeric or string variables. There are a number of advantages to converting categorical variables to factor variables. ... The only required argument is a vector of values which can be either string or numeric.

```{r}
train = read.csv('../raw_data/train (1).csv', stringsAsFactors=F)
dim(train)

test = read.csv('../raw_data/test (1).csv', stringsAsFactors = F)
test$Survived = NA
dim(test)

full = rbind(train, test) 
# check the structure
str(full)
head(full)
tail(full)
summary(full)
```

Missing value imputation

```{r}
# is there any Missing obesrvation
colSums(is.na(full))
# Empty data
colSums(full=='')

```

```{r}
table(full$Embarked)
# replace missing with mode
full$Embarked[full$Embarked==""]="S"
table(full$Embarked)
# Age and Cabin are missing too many for mode
```

Check the length and see how many varibles of them we can move to factor for our analysis

appy: Returns a vector or array or list of values obtained by applying a function to margins of an array or matrix.

Margin: a vector giving the subscripts which the function will be applied over. E.g., for a matrix 1 indicates rows, 2 indicates columns, c(1, 2) indicates rows and columns. Where X has named dimn ames, it can be a character vector selecting dimension names.

```{r}
apply(full,2, function(x) length(unique(x))) # number of unique values per column
```
nchar takes a character vector as an argument and returns a vector whose elements contain the sizes of the corresponding elements of x.
vector equivalent form of the if…else statement in R, the ifelse() function. So in R, you can't simply apply if-else to 
vector without 'for' statement. When using data$column, use the following:

AGE
```{r}

full$Age = ifelse(is.na(full$Age),
                          ave(full$Age, FUN = function(x) mean(x, na.rm = TRUE)),
                          full$Age)

# na.rm: a logical value indicating whether NA values should be stripped before the computation proceeds.
                     
```


FARE
```{r}

full$Fare = ifelse(is.na(full$Fare),
                     ave(full$Fare, FUN = function(x) mean(x, na.rm = TRUE)),
                     full$Fare)
```


CABIN
```{r}

# Process Cabin Column to show number of cabins passenger has
cabin = full$Cabin
n = length(cabin)
for(i in 1:n){
  if(nchar(cabin[i]) == 0){
    cabin[i] = 0
  } else{
    cabin[i] = length(strsplit(cabin[i]," ")[[1]])
  }
} 
table(cabin)
#looks good, use it for now
full$Cabin = cabin
str(full)

```

TITLE
```{r}
head(full$Name)

names = full$Name
title =  gsub("^.*, (.*?)\\..*$", "\\1", names)

table(title)
```

```{r}
n = length(title)
for(i in 1:n){
  if(title[i] == "Mlle"){
    title[i] = "Miss"
  }else if(title[i] == "Ms"){
    title[i] = "Miss"
  }else if(title[i] == "Mme"){
    title[i] = "Mrs"
  }else if(title[i] == "Lady"){
    title[i] = "Miss"
  }else if(title[i] == "Dona"){
    title[i] = "Mrs"
  }else if(title[i] == "Don"){
    title[i] = "Master"
  }else if(title[i] == "Capt"){
    title[i] = "crew"
  }else if(title[i] == "Col"){
    title[i] = "crew"
  }else if(title[i] == "Major"){
    title[i] = "crew"
  }else if(title[i] == "Rev"){
    title[i] = "crew"
  }else if(title[i] == "Dr"){
    title[i] = "crew"
  }else if(title[i] == "Jonkheer"){
    title[i] = "Master"
  }else if(title[i] == "Sir"){
    title[i] = "Master"
  }else if(title[i] == "the Countess"){
    title[i] = "Mrs"
  }
}

table(title)
full$title = title
```


FAMILY
```{r}
sib = full$SibSp
pat = full$Parch
family = sib + pat + 1
n = length(family)
for(i in 1:n){
  if(family[i] == 1){
    family[i] = "solo"
  }else if(family[i] >= 2 & family[i] < 5){
    family[i] = "small"
  }else{
    family[i] = "large"
  }
}

table(family)
full$family = family 
```


TICKET
```{r}
t_unique = rep(0, nrow(full)) #zero array used to construct new column
tickets = unique(full$Ticket)

str(tickets) #929 unique ticket values
str(full$Ticket) #1309
str(t_unique) #1309 zero vector
n = length(tickets)
for(i in 1:n){
  group = which(full$Ticket == tickets[i]) #integer row vector consisting of index numbers where this is TRUE
  m = length(group)
  
  for(k in 1:m){
    t_unique[group[k]] = m
  }
}


```

```{r}
t_size = c()
n = length(t_unique)
for(i in 1:n){
  if(t_unique[i] == 1){
    t_size[i] = "solo"
  }else if(t_unique[i] >= 2 & t_unique[i] < 5){
    t_size[i] = "small"
  }else{
    t_size[i] = "large" 
  }
}

str(t_size)  
table(t_size)
full$ticket_group = t_size

```



```{r}
str(full)
cols=c("Survived","Pclass","Sex","Embarked", "family", "title", "Cabin", "ticket_group")
for (i in cols){
  full[,i]=as.factor(full[,i])
}

contrasts(full$Pclass)

typeof(full$Age)

#ready to go for test
y = full[891:1309, c("Pclass","Sex","Embarked", "family", "title", "Age", "Cabin", "ticket_group", "Fare")]




X = full[1:891, c("Pclass","Sex","Embarked", "family", "title", "Age", "Cabin", "ticket_group", "Fare", "Survived")]

X$Age = scale(as.numeric(X$Age))
X$Fare = scale(as.numeric(X$Fare))

y$Age = scale(as.numeric(y$Age))
y$Fare = scale(as.numeric(y$Fare))


library(caTools)

split = sample.split(X$Survived, SplitRatio = 0.75)
training_set = subset(X, split == TRUE)
test_set = subset(X, split == FALSE)

```

```{r}

fit = glm(formula =  Survived ~ ., 
                 family = binomial,
                 data = training_set)

prob_pred = predict(fit, type = 'response', newdata = test_set)

head(prob_pred) # probability that dependent variable is equal to one
y_pred = ifelse(prob_pred > 0.5, 1, 0)
y_pred[1:10]
prob_pred[1:10]

# test_set[ , 10]
# Making the Confusion Matrix/accuracy
cm = as.matrix(table(Actual = test_set[ , 10], Predicted = y_pred))
paste("Accuracy = ", sum(diag(cm))/sum(cm))
```


```{r}
summary(fit)

```

Keep fare, ticket, Age, family

```{r}
fit = glm(formula =  Survived ~ Fare + ticket_group + Age + family + Sex + Pclass + title, 
                 family = binomial,
                 data = training_set)
#need to adjust test data for this
prob_pred = predict(fit, type = 'response', newdata = test_set)

head(prob_pred) # probability that dependent variable is equal to one
y_pred = ifelse(prob_pred > 0.5, 1, 0)
y_pred[1:10]

cm = as.matrix(table(Actual = test_set[ , 10], Predicted = y_pred))
paste("Accuracy = ", sum(diag(cm))/sum(cm))
```


Kernal SVM
```{r}
library(e1071)

fit = svm(formula =  Survived ~ ., data = training_set, 
          type = 'C-classification', 
          kernel = 'radial')
          
                 

prob_pred = predict(fit, type = 'response', newdata = test_set)

head(prob_pred) # don't need to write script for SVM 


# test_set[ , 10]
# Making the Confusion Matrix/accuracy
cm = as.matrix(table(Actual = test_set[ , 10], Predicted = prob_pred))
paste("Accuracy = ", sum(diag(cm))/sum(cm))
```



Naive Bayes

```{r}

fit = naiveBayes(x = training_set[-10],
                        y = training_set$Survived)

prob_pred = predict(fit, newdata = test_set) #binary output

cm = as.matrix(table(Actual = test_set[ , 10], Predicted = prob_pred))
paste("Accuracy = ", sum(diag(cm))/sum(cm))
```




NN needs work
```{r}
library(h2o)
h2o.init(nthreads = -1) #, max_mem_size = '2g', ip = "127.0.0.1", port = 54321
model = h2o.deeplearning(y = 'Exited',
                         training_frame = as.h2o(training_set),
                         activation = 'Rectifier',
                         hidden = c(5,5),
                         epochs = 100,
                         train_samples_per_iteration = -2)

```


















--------------------------------------------APPENDIX-----------------------------------------

```{r}
n = sum(cm) # number of instances
nc = nrow(cm) # number of classes
diag = diag(cm) # number of correctly classified instances per class 
rowsums = apply(cm, 1, sum) # number of instances per class
colsums = apply(cm, 2, sum) # number of predictions per class
p = rowsums / n # distribution of instances over the actual classes
q = colsums / n # distribution of instances over the predicted classes
cm
n
sum(diag)
accuracy = sum(diag) / n 
accuracy

#Precision is defined as the fraction of correct predictions for a certain class
precision = diag / colsums 

#recall is the fraction of instances of a class that were correctly predicted. 
recall = diag / rowsums 

#F-1 score is also commonly reported. It is defined as the harmonic mean (or a weighted average) of precision and recall.
f1 = 2 * precision * recall / (precision + recall) 

data.frame(precision, recall, f1) 
```


```{r}
ggplot(full[1:891,],aes(x = Cabin,fill=factor(Survived))) +
geom_bar() +
ggtitle("Cabin v/s Survival Rate")+
xlab("Cabin") +
ylab("Total Count") +
labs(fill = "Survived")  
```

```{r}
###Visualize P class which is the best proxy for Rich and Poor  

ggplot(full[1:891,],aes(x = Pclass,fill=factor(Survived))) +
geom_bar() +
ggtitle("Pclass v/s Survival Rate")+
xlab("Pclass") +
ylab("Total Count") +
labs(fill = "Survived")  
```


```{r}
ggplot(full[1:891,], aes(x = title, fill = Survived)) +
  geom_bar() +
  facet_wrap(~Pclass) + 
  ggtitle("3-way relationship of Title, Pclass, and Survival") +
  xlab("Title") +
  ylab("Total Count") +
  labs(fill = "Survived")
```

```{r}
# Visualize the 3-way relationship of sex, pclass, and survival
ggplot(full[1:891,], aes(x = Sex, fill = Survived)) +
geom_bar() + #bar chart
facet_wrap(~Pclass) + 
ggtitle("3D view of sex, pclass, and survival") +
xlab("Sex") +
ylab("Total Count") +
labs(fill = "Survived")
```


```{r}
# Visualize the 3-way relationship of sex, pclass, and survival
ggplot(full[1:891,], aes(x = Sex, fill = Survived)) +
geom_bar() + #bar chart
facet_wrap(~Pclass) + 
ggtitle("3D view of sex, pclass, and survival") +
xlab("Sex") +
ylab("Total Count") +
labs(fill = "Survived")
```